Dominance Regions for Imaginary $\bfg$-vectors

### 2021-10-04 dominance_regions.tex
Getting close to finishing the bijection we need.  Also some reorganization.

  \section{Bases From Compatible Pairs on Dyck Paths}
  \sayDR{This is a bit of a repetition of the above introduction to Dyck paths, but with slightly different goals it seemed that a modified exposition was more appropriate.}
  For $a_1,a_2\in\ZZ_{\ge0}$, choose a Dyck path $D$ in the lattice rectangle $[0,a_1]\times[0,a_2]$.
  When we wish to emphasize the upper right bounding corner we will write $D[a_1,a_2]$.
  Consider $D$ as the set of its edges taken as a totally ordered set along the natural labeling from $(0,0)$ to $(a_1,a_2)$.
  It may be convenient to identify $D$ with the ordered set $[1,a_1+a_2]$.
  Write $D=D_1\sqcup D_2$, where $D_1=\{h_1,\ldots,h_{a_1}\}$ is the set of horizontal edges and $D_2=\{v_1,\ldots,v_{a_2}\}$ is the set of vertical edges.
  For $h\in D_1$, define the \emph{height} $\hgt(h)$ as the number of $v\in D_2$ with $v<h$.
  For $v\in D_2$, define the \emph{depth} $\dpt(v)$ as the number of $h\in D_1$ with $h<v$.

@@

  Write $D^{max}=D^{max}[a_1,a_2]$ for the maximal Dyck path in the lattice rectangle $[0,a_1]\times[0,a_2]$, i.e. $D^{max}$ never crosses above the diagonal line joining $(0,0)$ to $(a_1,a_2)$ and any lattice point above $D^{max}$ also lies above the diagonal.

@@

  \begin{definition}
    \label{def:Dyck path mutations}
    Given a Dyck path $D=D[a_1,a_2]$ which is vertically-adapted to $c$, define a Dyck path $\mu_{1,c}(D)$ inside the lattice rectangle $[0,ca_1-a_2]\times[0,a_1]$ by replacing each horizontal edge of $D$, together with the $\ell$ vertical edges which immediately follow it, by $c-\ell$ horizontal edges followed by a vertical edge.
    This provides a canonical identification of the horizontal edges of $D$ with the vertical edges of $\mu_{1,c}(D)$, this identification is again denoted by $\mu_{1,c}$.

    Given a Dyck path $D=D[a_1,a_2]$ which is horizontally-adapted to $b$, define a Dyck path $\mu_{2,b}(D)$ inside the lattice rectangle $[0,a_2]\times[0,ba_2-a_1]$ by replacing each vertical edge of $D$, together with the $\ell$ horizontal edges which immediately precede it, by a horizontal edge followed by $b-\ell$ vertical edges.
    This provides a canonical identification of the vertical edges of $D$ with the horizontal edges of $\mu_{2,b}(D)$, this identification is again denoted by $\mu_{2,b}$.
  \end{definition}
  \begin{remark}
    Observe that $\mu_{1,a}$ and $\mu_{2,a}$ are inverse for any $a\ge1$ whenever they are defined.
  \end{remark}

@@

  \subsection{Compatibility}
  The results of this section can mostly be found in \cite{lee-li-zelevinsky,rupel}.
  However since we work in arbitrary Dyck paths instead of just the maximal ones, we recall and adapt the proofs as needed.

  For $e \le e'\in D$, write $ee'$ for the subpath of $D$ beginning with $e$ and ending with $e'$.
  %Take $\overline{e}e'$ and $e\overline{e}'$ to be the subpaths of $ee'$ with $e$ and $e'$ removed, respectively.
  When $e=e'$, we take $ee'$ to be the subpath consisting of only the edge $e$.
  For $e \le h\in D$ with $h\in D_1$, write $eh^{\!\uparrow}$ for the subpath $eh$ together with all vertical edges immediately following $h$.
  For $v \le e\in D$ with $v\in D_2$, write ${}^\leftarrow\!\!\! ve$ for the subpath $ve$ together with all horizontal edges immediately preceding $v$.
  \begin{definition}
    For $H\sqcup V\subset D$ with $H\subset D_1$ and $V\subset D_2$ together with a subpath $ee'\subset D$, define the \emph{horizontal shadow statistic}:
    \begin{equation}
      \label{eq:horizontal shadow statistic}
      f_H(ee')=-|ee' \cap D_2|+c|ee'\cap H|
    \end{equation}
    and the \emph{vertical shadow statistic}:
    \begin{equation}
      \label{eq:vertical shadow statistic}
      f_V(ee')=-|ee' \cap D_1|+b|ee'\cap V|.
    \end{equation}
  \end{definition}
  \begin{remark}
    Observe that the shadow statistics are additive with respect to concatenation of subpaths.
  \end{remark}

  \begin{lemma}
    Suppose the Dyck path $D$ is vertically-adapted to $c$ and let $D'=\mu_{1,c}(D)$ with $D'_2=\{v'_1,\ldots,v'_{a_1}\}$.
    For $H\subset D_1$, define $V'\subset D'_2$ by $V'=D'_2\setminus \mu_{1,c}(H)$.
    Then for any $h_i<h_j \in D_1$, we have $f_H(h_i h_j^{\!\uparrow})=-\vv{f}_{V'}({}^\leftarrow\!\!\! v'_i v'_j)$.
  \end{lemma}
  \begin{proof}
    By definition, we have 
    \begin{align*}
      -\vv{f}_{V'}({}^\leftarrow\!\!\! v'_iv'_j)&=|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_1|-c|{}^\leftarrow\!\!\! v'_i v'_j \cap V'|\\
      &=|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_1|-c\big|{}^\leftarrow\!\!\! v'_i v'_j \cap \big(D'_2\setminus \mu_{1,c}(H)\big)\big|\\
      &=|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_1|-c|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_2|+c\big|{}^\leftarrow\!\!\! v'_i v'_j \cap \mu_{1,c}(H)\big|.
    \end{align*}
    But, by construction of $\mu_{1,c}(D)$ and since $\mu_{2,c}(D')=D$, we have
    \[\big|{}^\leftarrow\!\!\! v'_i v'_j \cap \mu_{1,c}(H)\big|=|h_i h_j^{\!\uparrow} \cap H| \qquad \text{and} \qquad c|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_2|-|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_1|=|h_i h_j^{\!\uparrow} \cap D_2|\]
    giving
    \[-\vv{f}_{V'}({}^\leftarrow\!\!\! v'_iv'_j)=-|h_i h_j^{\!\uparrow} \cap D_2|+c|h_i h_j^{\!\uparrow} \cap H|=f_H(h_i h_j^{\!\uparrow})\]
    as desired.
  \end{proof}

  \begin{definition}
    Consider $H\sqcup V\subset D$ with $H\subset D_1$ and $V\subset D_2$.
    For $h\in H$ and $v\in V$ with $h<v$, call the path $hv$ \emph{compatible} if there exists an edge $e\in hv$ such that one of the following conditions is satisfied:
    \begin{equation}
      \label{eq:HGC}
      \tag{HGC} e\ne v \quad \text{and} \quad f_H(he)=0;
    \end{equation}
    \begin{equation}
      \label{eq:VGC}
      \tag{VGC} e\ne h \quad \text{and} \quad f_V(ev)=0.
    \end{equation}
    Call the subset $H\sqcup V$ \emph{compatible} if every path $hv$ with $h\in H$ and $v\in V$ is compatible.
    Write $\cC_D$ for the collection of all compatible subsets of $D$.
    Given a subset $H\subset D_1$, write $\cC_D(H)$ for the collection of all $V\subset D_2$ for which $H\sqcup V\in\cC_D$.
    Define $\cC_D(V)$ for $V\subset D_2$ similarly.
  \end{definition}
  \begin{remark}
    In the construction of the greedy basis, it is necessary to consider $D$ as a closed cycle by identifying $(0,0)$ with $(a_1,a_2)$ and allowing subpaths $ee'$ with $e'<e$ to ``wrap around'' $D$ in the obvious way.
    When we wish to emphasize that this condition should be imposed, we denote the Dyck path $D$ instead by $D_\WA$.
    Following \cite[Remark 2.21]{rupel}, for $D^{m,\varepsilon}$ and $\vv{\bar{D}}^{m,\varepsilon}$ the wrap around condition does not restrict compatibility and can thus safely be omitted.
  \end{remark}

  \begin{definition}
    \label{def:shadows}
    For $H\sqcup V\subset D$ with $H\subset D_1$ and $V\subset D_2$, define the \emph{shadow paths} as follows:
    \begin{itemize}
      \item For $h\in D_1$, write $D(h;H)$ (resp. $D_\WA(h;H)$) for the shortest subpath $he \subset D$ (resp. $he \subset D_\WA$) with $f_H(he)=0$ or, if no such subpath exists, set $D(h;H)=hv_{a_2}$ (resp. set $D_\WA(h;H)=D$).
      \item For $v\in D_2$, write $D(v;V)$ (resp. $D_\WA(v;V)$) for the shortest subpath $ev \subset D$ (resp. $ev \subset D_\WA$) with $f_V(ev)=0$ or, if no such subpath exists, set $D(v;V)=h_1 v$ (resp. set $D_\WA(v;V)=D$).
    \end{itemize}
    Write 
    \[D(H):=\bigcup_{h\in H} D(h;H) \qquad \text{and} \qquad D(V):=\bigcup_{v\in V} D(v;V).\]
    Define the \emph{shadow} of $H$ (resp. \emph{shadow} of $V$) as the set $\sh_D(H):=D(H)\cap D_2$ (resp. $\sh_D(V):=D(V)\cap D_1$) and the \emph{local shadow} of $h\in H$ (resp. of $v\in V$) as the set $\sh_D(h;H)=D(h;H)\cap D_2$ (resp. $\sh_D(v;V)=D(v;V)\cap D_1$).
    Define shadows and local shadows for $D_\WA$ similarly.
  \end{definition}

  \begin{lemma}
    \label{le:shadow statistics}
    Consider $H\sqcup V\subset D$ with $H\subset D_1$ and $V\subset D_2$.
    \begin{enumerate}
      \item For $h\in H$, take $he$ to be either the path $D(h;H)$ or $D_\WA(h;H)$.
        The following hold:
        \begin{enumerate}
          \item for any proper subpath $he'\subsetneq he$, we have $f_H(he')>0$;
          \item if $f_H(he)=0$, then for any proper subpath $e'e\subsetneq he$, we have $f_H(e'e)<0$;
          \item $e\in V$.
        \end{enumerate}
      \item For $v\in V$, take $ev$ to be either the path $D(v;V)$ or $D_\WA(v;V)$.
        The following hold:
        \begin{enumerate}
          \item for any proper subpath $e'v\subsetneq ev$, we have $f_V(e'v)>0$;
          \item if $f_V(ev)=0$, then for any proper subpath $ee'\subsetneq ev$, we have $f_V(ee')<0$;
          \item $e\in H$.
        \end{enumerate}
    \end{enumerate}
  \end{lemma}
  \begin{remark}
    Observe that the subset $V$ has no influence on the claim of (1) and the subset $H$ has no influence on the claim of (2).
  \end{remark}
  \begin{proof}
    We prove (1), the proof of (2) is similar.
    To see (a), note that as $e'$ traverses the path $he$ from $h$ to $e$, the value of $f_H(he')$ either increases by $c$, stays constant, or decreases by 1.
    Since $h\in H$, we have $f_H(hh)=c>0$ and so the value of $f_H(he')$ stays positive until it first becomes 0 when $e'=e$.
    Part (b) then follows from the additivity of $f_H$ and part (c) is the case when $e'=e$ here.
  \end{proof}

  \begin{corollary}
    Consider $H\sqcup V\subset D$ with $H\subset D_1$ and $V\subset D_2$.
    \begin{enumerate}
      \item For $h\in H$ and a vertical edge $v\in D(h;H)$, the condition \eqref{eq:HGC} is not satisfied for the path $hv$ but for any $V\in\cC_D(H)$ the condition \eqref{eq:VGC} is satisfied for the path $hv$.
        In particular, $D(v;V)$ is a proper subpath of $hv$ for any $V\in\cC_D(H)$.
      \item For $v\in V$ and a horizontal edge $h\in D(v;V)$, the condition \eqref{eq:VGC} is not satisfied for the pair $hv$ but for any $H\in\cC_D(V)$ the condition \eqref{eq:HGC} is satisfied for the path $hv$.
        In particular, $D(h;H)$ is a proper subpath of $hv$ for any $H\in\cC_D(V)$.
    \end{enumerate}
  \end{corollary}
  \begin{proof}
    We prove (1), the proof of (2) is similar.
    By Lemma~\ref{le:shadow statistics}, $f_H(he)>0$ for every proper subpath of $D(h;H)=hv$ and so the condition \eqref{eq:HGC} is not satisfied for the path $hv$, but then, since $V\in\cC_D(H)$, the condition \eqref{eq:VGC} must be satisfied for this path.
  \end{proof}

  \begin{lemma}\mbox{}
    \begin{enumerate}
      \item For $H\subset D_1$ and $h,h'\in H$, the local shadows $\sh_D(h;H)$ and $\sh_D(h';H)$ are either disjoint or one is contained in the other.
      \item For $V\subset D_2$ and $v,v'\in V$, the local shadows $\sh_D(v;V)$ and $\sh_D(v';V)$ are either disjoint or one is contained in the other.
    \end{enumerate}
  \end{lemma}
  \begin{proof}
    We prove (1), the proof of (2) is similar.
    For $h=h'$ or $D(h;H)=D$ or $D(h';H)=D$ there is nothing to show, so assume both shadow paths are proper subpaths of $D$.
    Suppose the intersection of $D(h;H)$ and $D(h';H)$ is nonempty, say with $h'\in D(h;H)=he$.
    If $e=v_{a_2}$, then $D(h';H)=h'v_{a_2}\subset D(h;H)$.
    Otherwise, denote by $e'\in D$ the edge immediately preceding $h'$ in $he$.
    If $e\in D(h';H)$, Lemma~\ref{le:shadow statistics} shows that $f_H(he)=f_H(he')+f_H(h'e)>0$, in contradiction to the definition of $D(h;H)$.
  \end{proof}

  \begin{definition}
    Consider $H\sqcup V\subset D$ with $H\subset D_1$ and $V\subset D_2$.
    \begin{enumerate}
      \item Define $\rsh_D(H)\subset\sh_D(H)$, the \emph{remote shadow} of $H$, by removing for each $d\in[1,a_1]$ the (up to) $c$ vertical edges of depth $d$ immediately following $h_d$.
      \item Define $\rsh_D(H)\subset\sh_D(H)$, the \emph{remote shadow} of $V$, by removing for each $\ell\in[1,a_2]$ the (up to) $b$ horizontal edges of height $\ell-1$ immediately preceding $v_\ell$.
    \end{enumerate}
  \end{definition}

  \begin{definition}
    Consider $H\sqcup V\subset D$ with $H\subset D_1$ and $V\subset D_2$.
    \begin{enumerate}
      \item For $0 < j,d \le a_1$ with $h_j\in H$, denote by $\rsh_D(H)_{j;d}$ the set of all $v\in\rsh_D(H)$ of depth $d$ such that $v\in\sh(h_j;H)$ and $h_j$ is the first horizontal edge before $v$ with this property.
        Define the \emph{local remote shadow} of the edge $h_j$ by $\rsh_D(h_j;H):=\bigsqcup_{d\in[1,a_1]} \rsh_D(H)_{j;d}$.
      \item For $0 < k \le a_2$ with $v_k\in V$ and $0\le \ell < a_2$, denote by $\rsh_D(V)_{k;\ell}$ the set of all $h\in\rsh_D(V)$ of height $\ell$ such that $h\in\sh(v_k;V)$ and $v_k$ is the first vertical edge after $h$ with this property.
        Define the \emph{local remote shadow} of the edge $v_k$ by $\rsh_D(v_k;V):=\bigsqcup_{\ell\in[0,a_2-1]} \rsh_D(V)_{k;\ell}$.
    \end{enumerate}
  \end{definition}

@@

  \begin{theorem}
    \label{th:dominant Dyck path recursion}
    For $a_1,a_2\in\ZZ_{\ge0}$ with $(-a_1+ba_2,-a_2)\in\cI$, we have $\mu_{1,c}\big(D^{dom}[a_1,a_2]\big)=\vv{D}^{dom}[ca_1-a_2,a_1]$ and $\mu_{2,b}\big(D^{dom}[a_1,a_2]\big)=\vv{D}^{dom}[a_2,ba_2-a_1]$.
  \end{theorem}
  \begin{proof}
    We prove the result for $\mu_{1,c}$ as the result for $\mu_{2,b}$ immediately follows from $\mu_{1,b}$ being its inverse.
    Write $D'$ for the Dyck path $\mu_{1,c}\big(D^{dom}[a_1,a_2]\big)$ and observe that by construction $D'$ has $ca_1-a_2$ horizontal edges and $a_1$ vertical edges.
    Write $v'_1,\ldots,v'_{a_1}$ for the vertical edges of $D'$ with $v'_r$ immediately preceded by $\ell_r$ horizontal edges.
    We have to show (i) that every lattice point on $D'$ stays below the lines 
    \[y=\frac{bc+\sqrt{bc(bc-4)}}{2c}x \qquad \text{and} \qquad y-a_1=\frac{bc-\sqrt{bc(bc-4)}}{2c}\big(x-(ca_1-a_2)\big)\]
    while (ii) any lattice point above $D'$ is also above one of these lines.

    If there exists $t$ so that $v'_t$ passes above the line $y=\frac{bc+\sqrt{bc(bc-4)}}{2c}x$, then
    \[ \frac{t}{\sum\limits_{r=1}^t \ell_r} > \frac{bc+\sqrt{bc(bc-4)}}{2c} \]
    or, rearranging, we have
    \[ \frac{\sum\limits_{r=1}^t \ell_r}{t} < \frac{2c}{bc+\sqrt{bc(bc-4)}} = \frac{bc-\sqrt{bc(bc-4)}}{2b} \]
    or
    \[ \frac{bc+\sqrt{bc(bc-4)}}{2b} < \frac{ct-\sum\limits_{r=1}^t \ell_r}{t} = \frac{\sum\limits_{r=1}^t (c-\ell_r)}{t}. \]
    But this states that the subpath of $D^{dom}[a_1,a_2]$ consisting of the first $t$ horizontal edges and all their trailing vertical edges passes above the line $y=\frac{bc+\sqrt{bc(bc-4)}}{2b}x$, contrary to the definition.

    If there exists $t$ so that $v'_t$ passes above the line $y-a_1=\frac{bc-\sqrt{bc(bc-4)}}{2c}\big(x-(ca_1-a_2)\big)$, then
    \[ \frac{bc-\sqrt{bc(bc-4)}}{2c} > \frac{a_1-t}{ca_1-a_2-\sum\limits_{r=1}^t \ell_r} \]
    or, rearranging, we have
    \[ \frac{ca_1-a_2-\sum\limits_{r=1}^t \ell_r}{a_1-t} > \frac{2c}{bc-\sqrt{bc(bc-4)}} = \frac{bc+\sqrt{bc(bc-4)}}{2b} \]
    or
    \[ \frac{bc-\sqrt{bc(bc-4)}}{2b} > \frac{ c(a_1-t) - \big(ca_1-a_2-\sum\limits_{r=1}^t \ell_r\big)}{a_1-t} = \frac{a_2-\sum\limits_{r=1}^t (c-\ell_r)}{a_1-t}. \]
    But this states that the subpath of $D^{dom}[a_1,a_2]$ consisting of the first $t$ horizontal edges and all their trailing vertical edges passes above the line $y-a_2=\frac{bc-\sqrt{bc(bc-4)}}{2b}(x-a_1)$, contrary to the definition.

@@

  This can be made more precise as follows.
  \begin{lemma}
    For $s>0$, choose $m\ge2$ minimal so that $s < u_{m,-}$
  \end{lemma}

@@

>>>>

  \section{Bases From Compatible Pairs on Dyck Paths}
  \sayDR{This is a bit of a repetition of the above introduction to Dyck paths, but with slightly different goals it seemed that a modified exposition was more appropriate.}
  For $a_1,a_2\in\ZZ_{\ge0}$, choose a Dyck path $D$ in the lattice rectangle $[0,a_1]\times[0,a_2]$.
  When we wish to emphasize the upper right bounding corner we will write $D[a_1,a_2]$.
  Consider $D$ as the totally ordered set of its edges along the natural labeling from $(0,0)$ to $(a_1,a_2)$.
  It may be convenient to identify $D$ with the ordered set $[1,a_1+a_2]$.
  Write $D=D_1\sqcup D_2$, where $D_1=\{h_1,\ldots,h_{a_1}\}$ is the set of horizontal edges and $D_2=\{v_1,\ldots,v_{a_2}\}$ is the set of vertical edges.
  Define the \emph{height}, $\hgt(h)$ for $h\in D_1$, as one more than the number of $v\in D_2$ with $v<h$.
  Define the \emph{depth}, $\dpt(v)$ for $v\in D_2$, as the number of $h\in D_1$ with $h<v$.
  \begin{remark}
    The seemingly different conventions for computing height and depth provide uniformity later.
    Indeed, the vertical edges immediately following $h_d$ have depth $d$, while this convention allows the horizontal edges immediately preceding $v_\ell$ to similarly be said to have height $\ell$.
  \end{remark}

@@

  \subsection{Structure of Dyck Paths}

  Write $D^{max}=D^{max}[a_1,a_2]$ for the maximal Dyck path in the lattice rectangle $[0,a_1]\times[0,a_2]$, i.e. $D^{max}$ never crosses above the diagonal line joining $(0,0)$ to $(a_1,a_2)$ and any lattice point above $D^{max}$ also lies above the diagonal.
  Observe that $\bar{D}^{max}[a_1,a_2]=D^{max}[a_2,a_1]$.

@@

  \begin{definition}
    \label{def:Dyck path mutations}
    Given a Dyck path $D=D[a_1,a_2]$ which is vertically-adapted to $c$, define a Dyck path $\mu_{1,c}(D)$ inside the lattice rectangle $[0,ca_1-a_2]\times[0,a_1]$ by replacing each horizontal edge of $D$, together with the $r$ vertical edges which immediately follow it, by $c-r$ horizontal edges followed by a vertical edge.
    This provides a canonical identification of the horizontal edges of $D$ with the vertical edges of $\mu_{1,c}(D)$, this identification is again denoted by $\mu_{1,c}$.

    Given a Dyck path $D=D[a_1,a_2]$ which is horizontally-adapted to $b$, define a Dyck path $\mu_{2,b}(D)$ inside the lattice rectangle $[0,a_2]\times[0,ba_2-a_1]$ by replacing each vertical edge of $D$, together with the $r$ horizontal edges which immediately precede it, by a horizontal edge followed by $b-r$ vertical edges.
    This provides a canonical identification of the vertical edges of $D$ with the horizontal edges of $\mu_{2,b}(D)$, this identification is again denoted by $\mu_{2,b}$.
  \end{definition}
  \begin{remark}
    Observe that $\mu_{1,a}$ and $\mu_{2,a}$ are inverse for any $a\ge1$ whenever they are defined.
  \end{remark}

@@

@@

  \begin{theorem}
    \label{th:dominant Dyck path recursion}
    For $a_1,a_2\in\ZZ_{\ge0}$ with $(-a_1+ba_2,-a_2)\in\cI$, we have $\mu_{1,c}\big(D^{dom}[a_1,a_2]\big)=\vv{D}^{dom}[ca_1-a_2,a_1]$ and $\mu_{2,b}\big(D^{dom}[a_1,a_2]\big)=\vv{D}^{dom}[a_2,ba_2-a_1]$.
  \end{theorem}
  \begin{proof}
    We prove the result for $\mu_{1,c}$ as the result for $\mu_{2,b}$ immediately follows from $\mu_{1,b}$ being its inverse.
    Write $D'$ for the Dyck path $\mu_{1,c}\big(D^{dom}[a_1,a_2]\big)$ and observe that by construction $D'$ has $ca_1-a_2$ horizontal edges and $a_1$ vertical edges.
    Write $v'_1,\ldots,v'_{a_1}$ for the vertical edges of $D'$ with $v'_k$ immediately preceded by $r_k$ horizontal edges.
    We have to show (i) that every lattice point on $D'$ stays below the lines 
    \[y=\frac{bc+\sqrt{bc(bc-4)}}{2c}x \qquad \text{and} \qquad y-a_1=\frac{bc-\sqrt{bc(bc-4)}}{2c}\big(x-(ca_1-a_2)\big)\]
    while (ii) any lattice point above $D'$ is also above one of these lines.

    If there exists $t$ so that $v'_t$ passes above the line $y=\frac{bc+\sqrt{bc(bc-4)}}{2c}x$, then
    \[ \frac{t}{\sum\limits_{k=1}^t r_k} > \frac{bc+\sqrt{bc(bc-4)}}{2c} \]
    or, rearranging, we have
    \[ \frac{\sum\limits_{k=1}^t r_k}{t} < \frac{2c}{bc+\sqrt{bc(bc-4)}} = \frac{bc-\sqrt{bc(bc-4)}}{2b} \]
    or
    \[ \frac{bc+\sqrt{bc(bc-4)}}{2b} < \frac{ct-\sum\limits_{k=1}^t r_k}{t} = \frac{\sum\limits_{k=1}^t (c-r_k)}{t}. \]
    But this states that the subpath of $D^{dom}[a_1,a_2]$ consisting of the first $t$ horizontal edges and all their trailing vertical edges passes above the line $y=\frac{bc+\sqrt{bc(bc-4)}}{2b}x$, contrary to the definition.

    If there exists $t$ so that $v'_t$ passes above the line $y-a_1=\frac{bc-\sqrt{bc(bc-4)}}{2c}\big(x-(ca_1-a_2)\big)$, then
    \[ \frac{bc-\sqrt{bc(bc-4)}}{2c} > \frac{a_1-t}{ca_1-a_2-\sum\limits_{k=1}^t r_k} \]
    or, rearranging, we have
    \[ \frac{ca_1-a_2-\sum\limits_{k=1}^t r_k}{a_1-t} > \frac{2c}{bc-\sqrt{bc(bc-4)}} = \frac{bc+\sqrt{bc(bc-4)}}{2b} \]
    or
    \[ \frac{bc-\sqrt{bc(bc-4)}}{2b} > \frac{ c(a_1-t) - \big(ca_1-a_2-\sum\limits_{k=1}^t r_k\big)}{a_1-t} = \frac{a_2-\sum\limits_{k=1}^t (c-r_k)}{a_1-t}. \]
    But this states that the subpath of $D^{dom}[a_1,a_2]$ consisting of the first $t$ horizontal edges and all their trailing vertical edges passes above the line $y-a_2=\frac{bc-\sqrt{bc(bc-4)}}{2b}(x-a_1)$, contrary to the definition.

@@

@@

  \subsection{Compatibility}
  The results of this section can mostly be found in \cite{lee-li-zelevinsky,rupel}.
  However since we work in arbitrary Dyck paths instead of just the maximal ones, some results and proofs will need to be adapted.

  For $e \le e'\in D$, write $ee'$ for the subpath of $D$ beginning with $e$ and ending with $e'$.
  When $e=e'$, we take $ee'$ to be the subpath consisting of only the edge $e$.
  \begin{definition}
    For $H\sqcup V\subset D$ with $H\subset D_1$ and $V\subset D_2$ together with a subpath $ee'\subset D$, define the \emph{horizontal shadow statistic}:
    \begin{equation}
      \label{eq:horizontal shadow statistic}
      f_H(ee')=-|ee' \cap D_2|+c|ee'\cap H|
    \end{equation}
    and the \emph{vertical shadow statistic}:
    \begin{equation}
      \label{eq:vertical shadow statistic}
      f_V(ee')=-|ee' \cap D_1|+b|ee'\cap V|.
    \end{equation}
  \end{definition}
  \begin{remark}
    Observe that the shadow statistics are additive with respect to concatenation of subpaths.
  \end{remark}

  A key notion from \cite{lee-li-zelevinsky} is that of ``compatible subsets of a maximal Dyck path'', this construction immediately generalizes to arbitrary Dyck paths as follows.
  \begin{definition}
    Consider $H\sqcup V\subset D$ with $H\subset D_1$ and $V\subset D_2$.
    For $h\in H$ and $v\in V$ with $h<v$, call the path $hv$ \emph{compatible} if there exists an edge $e\in hv$ such that one of the following conditions is satisfied:
    \begin{equation}
      \label{eq:HGC}
      \tag{HGC} e\ne v \quad \text{and} \quad f_H(he)=0;
    \end{equation}
    \begin{equation}
      \label{eq:VGC}
      \tag{VGC} e\ne h \quad \text{and} \quad f_V(ev)=0.
    \end{equation}
    Call the subset $H\sqcup V$ \emph{compatible} if every path $hv$ with $h\in H$ and $v\in V$ is compatible.
    Write $\cC_D$ for the collection of all compatible subsets of $D$.
    Given a subset $H\subset D_1$, write $\cC_D(H)$ for the collection of all $V\subset D_2$ for which $H\sqcup V\in\cC_D$.
    Define $\cC_D(V)$ for $V\subset D_2$ similarly.
  \end{definition}
  \begin{remark}
    In the construction of the greedy basis, it is necessary to consider $D$ as a closed cycle by identifying $(0,0)$ with $(a_1,a_2)$ and allowing subpaths $ee'$ with $e'<e$ to ``wrap around'' $D$ in the obvious way.
    When we wish to emphasize that this condition should be imposed, we denote the Dyck path $D$ instead by $D_\WA$.
    Following \cite[Remark 2.21]{rupel}, for $D^{m,\varepsilon}$ and $\vv{\bar{D}}^{m,\varepsilon}$ the wrap around condition does not restrict compatibility and can thus safely be omitted.
  \end{remark}

  The crucial result from \cite{lee-li-zelevinsky} can be stated as follows: if a Dyck path $D$ is vertically-adapted to $c$ (resp. horizontally adapted to $b$), then there is a bijection between certain compatible subsets of $D$ and certain compatible subsets of $\mu_{1,c}(D)$ (resp. of $\mu_{2,b}(D)$).
  The main step in the proof that needs to be adapted to the general situation is the following.
  For $e \le h\in D$ with $h=h_d\in D_1$, write $eh^{\!\uparrow}$ for the subpath $eh$ together with all vertical edges of depth $d$ immediately following $h$.
  For $v \le e\in D$ with $v=v_\ell\in D_2$, write ${}^\leftarrow\!\!\! ve$ for the subpath $ve$ together with all horizontal edges of height $\ell$ immediately preceding $v$.
  \begin{lemma}\mbox{}
    \label{le:shadow statistic recursion}
    \begin{enumerate}
      \item Suppose the Dyck path $D$ is vertically-adapted to $c$ and let $D'=\mu_{1,c}(D)$ with $D'_2=\{v'_1,\ldots,v'_{a_1}\}$.
        For $H\subset D_1$, define $V'\subset D'_2$ by $V'=D'_2\setminus \mu_{1,c}(H)$.
        Then for any $h_i \le h_j \in D_1$, we have $f_H(h_i h_j^{\!\uparrow})=-\vv{f}_{V'}({}^\leftarrow\!\!\! v'_i v'_j)$.
      \item Suppose the Dyck path $D$ is horizontally-adapted to $b$ and let $D''=\mu_{2,b}(D)$ with $D''_1=\{h''_1,\ldots,h''_{a_2}\}$.
        For $V\subset D_2$, define $H''\subset D''_1$ by $H''=D''_1\setminus \mu_{2,b}(V)$.
        Then for any $v_i \le v_j \in D_2$, we have $f_V({}^\leftarrow\!\!\! v_i v_j)=-\vv{f}_{H''}(h''_i {h''_j}^{\!\uparrow})$.
    \end{enumerate}
  \end{lemma}
  \begin{proof}
    Claim (2) is immediate from claim (1). 
    By definition, we have 
    \begin{align*}
      -\vv{f}_{V'}({}^\leftarrow\!\!\! v'_iv'_j)&=|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_1|-c|{}^\leftarrow\!\!\! v'_i v'_j \cap V'|\\
      &=|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_1|-c\big|{}^\leftarrow\!\!\! v'_i v'_j \cap \big(D'_2\setminus \mu_{1,c}(H)\big)\big|\\
      &=|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_1|-c|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_2|+c\big|{}^\leftarrow\!\!\! v'_i v'_j \cap \mu_{1,c}(H)\big|.
    \end{align*}
    But, by construction of $\mu_{1,c}(D)$ and since $\mu_{2,c}(D')=D$, we have
    \[\big|{}^\leftarrow\!\!\! v'_i v'_j \cap \mu_{1,c}(H)\big|=|h_i h_j^{\!\uparrow} \cap H| \qquad \text{and} \qquad c|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_2|-|{}^\leftarrow\!\!\! v'_i v'_j \cap D'_1|=|h_i h_j^{\!\uparrow} \cap D_2|\]
    giving
    \[-\vv{f}_{V'}({}^\leftarrow\!\!\! v'_iv'_j)=-|h_i h_j^{\!\uparrow} \cap D_2|+c|h_i h_j^{\!\uparrow} \cap H|=f_H(h_i h_j^{\!\uparrow})\]
    as desired.
  \end{proof}

  The desired bijection of compatible subsets is constructed in terms of certain subpaths determined from each compatible subset.
  \begin{definition}
    \label{def:shadows}
    For $H\sqcup V\subset D$ with $H\subset D_1$ and $V\subset D_2$, define the \emph{shadow paths} as follows:
    \begin{itemize}
      \item For $h\in D_1$, write $D(h;H)$ for the shortest subpath $he \subset D$ with $f_H(he)=0$ or, if no such subpath exists, set $D(h;H)=hv_{a_2}$.
      \item For $v\in D_2$, write $D(v;V)$ for the shortest subpath $ev \subset D$ with $f_V(ev)=0$ or, if no such subpath exists, set $D(v;V)=h_1 v$.
    \end{itemize}
    Write 
    \[D(H):=\bigcup_{h\in H} D(h;H) \qquad \text{and} \qquad D(V):=\bigcup_{v\in V} D(v;V).\]
    Define the \emph{shadow} of $H$ (resp. the \emph{shadow} of $V$) as the set $\sh_D(H):=D(H)\cap D_2$ (resp.~$\sh_D(V):=D(V)\cap D_1$) and the \emph{local shadow} of $h\in H$ (resp. of $v\in V$) as the set $\sh_D(h;H)=D(h;H)\cap D_2$ (resp. $\sh_D(v;V)=D(v;V)\cap D_1$).
  \end{definition}

  The proofs of the following results are identical to the proof of the analogous results from \cite{lee-li-zelevinsky,rupel} for maximal Dyck paths.
  \begin{lemma}
    \label{le:shadow statistics}
    Consider $H\sqcup V\subset D$ with $H\subset D_1$ and $V\subset D_2$.
    \begin{enumerate}
      \item For $h\in H$, take $he$ to be the path $D(h;H)$.
        The following hold:
        \begin{enumerate}
          \item for any proper subpath $he'\subsetneq he$, we have $f_H(he')>0$;
          \item if $f_H(he)=0$, then for any proper subpath $e'e\subsetneq he$, we have $f_H(e'e)<0$;
          \item $e\in V$.
        \end{enumerate}
      \item For $v\in V$, take $ev$ to be the path $D(v;V)$.
        The following hold:
        \begin{enumerate}
          \item for any proper subpath $e'v\subsetneq ev$, we have $f_V(e'v)>0$;
          \item if $f_V(ev)=0$, then for any proper subpath $ee'\subsetneq ev$, we have $f_V(ee')<0$;
          \item $e\in H$.
        \end{enumerate}
    \end{enumerate}
  \end{lemma}
  \begin{remark}
    Observe that the subset $V$ has no influence on the claim of (1) and the subset $H$ has no influence on the claim of (2).
  \end{remark}
